Behavior Driven Development in Haskell

### 2012-05-24 paper.tex
writing problems.

In this paper, we consider how put BDD into practice in Haskell.
We brought BDD methods from other programming language communities and
integrated them with a Haskell-original test method.
In particular, we introduced {\tt doctest}~\cite{doctest} and
integrated with QuickCheck~\cite{quickcheck}. 
We also implemented {\tt hspec} inspired by rspec~\cite{rspec} 
and integrated with QuickCheck.

@@

\begin{description}
\item[developer] programmers who develop libraries
\item[users] programmers who use libraries
\item[unit test] tests for individual programs or libraries.
\item[example] Typical usage of functions.
\item[property] Invariants which functions have.
\item[behavior] How do functions behave for users. Behavior can be shown by using both examples and properties.
\item[design] TBD
\item[documentation] Explanation of functions.
\end{description}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observation}

We insist that examples and properties, which are useful for users, should be
documented.
So, let us first observe Haddock~\cite{haddock} (version 2.7.2 or earlier), the documentation system of Haskell. 
Haddock defines annotation syntax in Haskell's comment.

Haddock provides code block annotations indicated with ({\tt >}).
For example, the {\tt container} packages has the following code block annotation:

@@

@@

TBD: Talk about test-framework. it integrates HUnit and QuickCheck but not 
suitable for design.

>>>>

In this paper, we consider how to put BDD into practice in Haskell.
We brought BDD methods from other programming language communities and
integrated them with a Haskell-original test method.
In particular, we introduced {\tt doctest}~\cite{doctest} and
integrated with QuickCheck~\cite{quickcheck}. 
We also implemented {\tt hspec} inspired by rspec~\cite{rspec} 
and integrated with QuickCheck.

@@

\begin{description}
\item[developer] programmers who develop libraries
\item[users] programmers who use libraries
%%\item[unit test] tests for individual programs or libraries.
\item[example] Typical usage of functions.
\item[property] Invariants which functions have.
\item[behavior] How do functions behave for users. Behavior can be shown by using both examples and properties.
\item[documentation] Explanation of functions.
\end{description}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problems}

In Haskell community, HUnit and QuickCheck are commonly used %% what about SmallCheck?
and the {\tt test-framework} package integrates them. 
We can categorize the {\tt test-framework} package into
a tool of TDD.
Before our work, there is no BDD tool commonly used in Haskell community.

BDD tools used in other programming language communities are
mainly based on example test. When we bring them to
Haskell community, we should integrate them with properties test.

We can categorize behaviors into two types: visible behaviors
and internal behaviors.
Visible behaviors are those which are useful for for both developers and users.
By contrast, internal behaviors are those which are usefull only for developers.
Behavior for a corner case registered in a bug track system is
a good example of internal behaviors.

BDD tools should be treat both of them and provide functionality
to disclose visible behaviors as documentation for users.

TBD: doctest and comments

The standard documentation mechanism for Haskell is
Haddock~\cite{haddock}.
Haddock defines annotation syntax in Haskell's comment.
Haddock provides code block annotations indicated with ({\tt >}).
For example, the {\tt container} packages has the following code block annotation:

@@

Haddock version 2.7.2 or earlier does not provide annotation syntax 
for examples and properties.

@@

Here are summary of the problems:

\begin{itemize}
\item Haskell community does not have common syntax to describe
behaviors in Haddock.
\item Haskell community does not have a BBD tool for visible behaviors.
\item Haskell community does not have a BBD tool for internal behaviors.
\end{itemize}
